infer: An R package for tidyverse-friendly statistical inference

### 2021-06-24 figs/paper/paper.tex
ben's edits

        \author[1, 2]{Simon Couch}
          \author[3]{Andrew Bray}
          \author[4]{Chester Ismay}
          \author[5]{Evgeni Chasnovski}
          \author[6]{Ben Baumer}
          \author[2, 7]{Mine Çetinkaya-Rundel}

@@

\texttt{infer} implements an expressive grammar to perform statistical
inference that coheres with the \texttt{tidyverse} design framework
(Wickham et al., 2019). Rather than providing methods for specific
statistical tests, this package consolidates the principles that are
shared among common hypothesis tests into a set of four main verbs
(functions), supplemented with many utilities to visualize and extract
value from their outputs.

@@

infer is an R package for randomization-based hypothesis testing,
naturalizing an intuitive understanding of statistical inference via a
unified and expressive grammar. Four functions provide functionality
encompassing a large swath of basic frequentist statistical inference,
abstracting away details of specific tests and shifting the focus of the
analyst to the observed data and the processes that generated it. Such a
grammar lends itself to applications in teaching, data pedagogy
research, applied scientific research, and advanced predictive modeling.
For one, the principled approach of the package has made it especially
well-fit for use in teaching introductory statistics (Baumer, Bray,
Çetinkaya-Rundel, \& Hardin, 2020; Ismay \& Kim, 2019) and research in
data pedagogy (Çetinkaya-Rundel \& Ellison, 2021; Fergusson \&
Pfannkuch, 2021; Loy, 2021). Further, the package has already seen usage
in a number of published scientific applications (Ask et al., 2021;
Fallon \& Hinds, 2021; McLean et al., 2021). Finally, the package
integrates with the greater tidymodels collection of packages, a
burgeoning software ecosystem for tidyverse-aligned predictive modeling
used across many modern research and industrial applications (Kuhn \&
Wickham, 2020). To date, the package has been downloaded more than
400,000 times.

@@

\begin{itemize}
\tightlist
\item
  \texttt{specify()} allows the analyst to specify the variable, or
  relationship between variables, that they're interested in.
\item
  \texttt{hypothesize()} allows the analyst to declare the null
  hypothesis.
\item
  \texttt{generate()} allows the analyst to generate data reflecting the
  null hypothesis.
\item
  \texttt{calculate()} allows the analyst to calculate summary
  statistics, either from

@@

Beyond this, the package offers

@@

\texttt{broom} and \texttt{parameters} convert model objects to unified
output formats, though do not provide methods for fitting models,
describing null distributions, or calculating summary statistics from
tabular data. \texttt{statsexpressions}, and adjacent packages in the
\texttt{easystats} ecosystem, implement wrappers with consistent
interfaces for theory-based hypothesis tests. At a higher level, the
structure of each of these packages is defined by model types, where
each model type has its own associated function and/or model object. In
contrast, \texttt{infer} is structured around four functions, situating
statistics and model types within a more abstracted grammar.
\texttt{infer}'s initial release on the Comprehensive R Archive Network
predated that of each of the packages mentioned above, with the
exception of \texttt{broom} (R Core Team, 2021).

@@

We acknowledge contributions from Albert Kim, Jo Hardin, Jay Lee, Amelia
McNamara, Nick Solomon, and Richie Cotton.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-ask2021per}{}%
Ask, A. V., Jenssen, B. M., Tartu, S., Angelier, F., Chastel, O., \&
Gabrielsen, G. W. (2021). Per-and polyfluoroalkyl substances are
positively associated with thyroid hormones in an arctic seabird.
\emph{Environmental Toxicology and Chemistry}, \emph{40}(3), 820--831.

\leavevmode\hypertarget{ref-baumer2020teaching}{}%
Baumer, B. S., Bray, A. P., Çetinkaya-Rundel, M., \& Hardin, J. S.
(2020). Teaching introductory statistics with DataCamp. \emph{Journal of
Statistics Education}, \emph{28}(1), 89--97.

\leavevmode\hypertarget{ref-cetinkaya2021fresh}{}%
Çetinkaya-Rundel, M., \& Ellison, V. (2021). A fresh look at
introductory data science. \emph{Journal of Statistics and Data Science
Education}, \emph{29}(sup1), S16--S26.

\leavevmode\hypertarget{ref-fallon2021single}{}%
Fallon, M. E., \& Hinds, M. T. (2021). Single cell morphological metrics
and cytoskeletal alignment regulate VCAM-1 protein expression.
\emph{Biochemical and Biophysical Research Communications}, \emph{555},
160--167.

\leavevmode\hypertarget{ref-fergusson2021introducing}{}%
Fergusson, A., \& Pfannkuch, M. (2021). Introducing teachers who use
GUI-driven tools for the randomization test to code-driven tools.
\emph{Mathematical Thinking and Learning}, 1--21.

\leavevmode\hypertarget{ref-ismay2019statistical}{}%
Ismay, C., \& Kim, A. Y. (2019). \emph{Statistical inference via data
science: A ModernDive into {R} and the tidyverse}. CRC Press.

\leavevmode\hypertarget{ref-kuhn2020tidymodels}{}%
Kuhn, M., \& Wickham, H. (2020). Tidymodels: A collection of packages
for modeling and machine learning using tidyverse principles.
\emph{Boston, MA, USA.}

\leavevmode\hypertarget{ref-loy2021bringing}{}%
Loy, A. (2021). Bringing visual inference to the classroom.
\emph{Journal of Statistics and Data Science Education}, 1--12.

\leavevmode\hypertarget{ref-mclean2021controlled}{}%
McLean, Z. L., Appleby, S. J., Fermin, L. M., Henderson, H. V., Wei, J.,
Wells, D. N., \& Oback, B. (2021). Controlled cytoplast arrest and
morula aggregation enhance development, cryoresilience, and in vivo
survival of cloned sheep embryos. \emph{Cellular Reprogramming},
\emph{23}(1), 14--25.

\leavevmode\hypertarget{ref-CRAN}{}%
R Core Team. (2021). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-wickham2019welcome}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., et al. (2019). Welcome to the tidyverse.
\emph{Journal of Open Source Software}, \emph{4}(43), 1686.

>>>>

        \author[1, 2]{Simon P. Couch}
          \author[3]{Andrew P. Bray}
          \author[4]{Chester Ismay}
          \author[5]{Evgeni Chasnovski}
          \author[6]{Benjamin S. Baumer}
          \author[2, 7]{Mine Çetinkaya-Rundel}

@@

\texttt{infer} implements an expressive grammar to perform statistical
inference that adheres to the \texttt{tidyverse} design framework
(Wickham et al., 2019). Rather than providing methods for specific
statistical tests, this package consolidates the principles that are
shared among common hypothesis tests into a set of four main verbs
(functions), supplemented with many utilities to visualize and extract
value from their outputs.

@@

\texttt{infer} is an R package for randomization-based hypothesis
testing, naturalizing an intuitive understanding of statistical
inference via a unified and expressive grammar. Four functions provide
functionality encompassing a large swath of basic frequentist
statistical inference, abstracting away details of specific tests and
shifting the focus of the analyst to the observed data and the processes
that generated it. Such a grammar lends itself to applications in
teaching, data pedagogy research, applied scientific research, and
advanced predictive modeling. For one, the principled approach of the
package has made it an especially good fit for teaching introductory
statistics (Baumer, Bray, Çetinkaya-Rundel, \& Hardin, 2020; Ismay \&
Kim, 2019) and research in data pedagogy (Çetinkaya-Rundel \& Ellison,
2021; Fergusson \& Pfannkuch, 2021; Loy, 2021). Further, the package has
already seen usage in a number of published scientific applications (Ask
et al., 2021; Fallon \& Hinds, 2021; McLean et al., 2021). Finally, the
package integrates with the greater tidymodels collection of packages, a
burgeoning software ecosystem for tidyverse-aligned predictive modeling
used across many modern research and industrial applications (Kuhn \&
Wickham, 2020). To date, the package has been downloaded more than
400,000 times.

@@

\begin{itemize}
\tightlist
\item
  \texttt{specify()} allows the analyst to specify the variable, or
  relationship between variables, that they are interested in.
\item
  \texttt{hypothesize()} allows the analyst to declare the null
  hypothesis.
\item
  \texttt{generate()} allows the analyst to generate data reflecting the
  null hypothesis.
\item
  \texttt{calculate()} allows the analyst to calculate summary
  statistics, either from

@@

Beyond this, the package offers:

@@

\texttt{broom} and \texttt{parameters} convert model objects to unified
output formats, though do not provide methods for fitting models,
describing null distributions, or calculating summary statistics from
tabular data. \texttt{statsexpressions}, and adjacent packages in the
\texttt{easystats} ecosystem, implement wrappers with consistent
interfaces for theory-based hypothesis tests. At a higher level, the
structure of each of these packages is defined by model types, where
each model type has its own associated function and/or model object. In
contrast, \texttt{infer} is structured around four functions, situating
statistics and model types within a more abstracted grammar.
\texttt{infer}'s initial release on the Comprehensive R Archive Network
predated that of each of the packages mentioned above, with the
exception of \texttt{broom}, a package situated in the same
\texttt{tidymodels} ecosystem (R Core Team, 2021).

@@

We acknowledge contributions from Albert Y. Kim, Jo Hardin, Jay Lee,
Amelia McNamara, Nick Solomon, and Richie Cotton.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-ask2021per}{}%
Ask, A. V., Jenssen, B. M., Tartu, S., Angelier, F., Chastel, O., \&
Gabrielsen, G. W. (2021). Per-and polyfluoroalkyl substances are
positively associated with thyroid hormones in an arctic seabird.
\emph{Environmental Toxicology and Chemistry}, \emph{40}(3), 820--831.
doi:\href{https://doi.org/10.1002/etc.4978}{10.1002/etc.4978}

\leavevmode\hypertarget{ref-baumer2020teaching}{}%
Baumer, B. S., Bray, A. P., Çetinkaya-Rundel, M., \& Hardin, J. S.
(2020). Teaching introductory statistics with DataCamp. \emph{Journal of
Statistics Education}, \emph{28}(1), 89--97.
doi:\href{https://doi.org/10.1080/10691898.2020.1730734}{10.1080/10691898.2020.1730734}

\leavevmode\hypertarget{ref-cetinkaya2021fresh}{}%
Çetinkaya-Rundel, M., \& Ellison, V. (2021). A fresh look at
introductory data science. \emph{Journal of Statistics and Data Science
Education}, \emph{29}(sup1), S16--S26.
doi:\href{https://doi.org/10.1080/10691898.2020.1804497}{10.1080/10691898.2020.1804497}

\leavevmode\hypertarget{ref-fallon2021single}{}%
Fallon, M. E., \& Hinds, M. T. (2021). Single cell morphological metrics
and cytoskeletal alignment regulate VCAM-1 protein expression.
\emph{Biochemical and Biophysical Research Communications}, \emph{555},
160--167.
doi:\href{https://doi.org/10.1016/j.bbrc.2021.03.129}{10.1016/j.bbrc.2021.03.129}

\leavevmode\hypertarget{ref-fergusson2021introducing}{}%
Fergusson, A., \& Pfannkuch, M. (2021). Introducing teachers who use
GUI-driven tools for the randomization test to code-driven tools.
\emph{Mathematical Thinking and Learning}, 1--21.
doi:\href{https://doi.org/10.1080/10986065.2021.1922856}{10.1080/10986065.2021.1922856}

\leavevmode\hypertarget{ref-ismay2019statistical}{}%
Ismay, C., \& Kim, A. Y. (2019). \emph{Statistical inference via data
science: A ModernDive into {R} and the tidyverse}. CRC Press.
doi:\href{https://doi.org/10.1080/00224065.2020.1848366}{10.1080/00224065.2020.1848366}

\leavevmode\hypertarget{ref-kuhn2020tidymodels}{}%
Kuhn, M., \& Wickham, H. (2020). Tidymodels: A collection of packages
for modeling and machine learning using tidyverse principles.
\emph{Boston, MA, USA.} Retrieved from \url{https://tidymodels.org}

\leavevmode\hypertarget{ref-loy2021bringing}{}%
Loy, A. (2021). Bringing visual inference to the classroom.
\emph{Journal of Statistics and Data Science Education}, 1--12.
doi:\href{https://doi.org/10.1080/26939169.2021.1920866}{10.1080/26939169.2021.1920866}

\leavevmode\hypertarget{ref-mclean2021controlled}{}%
McLean, Z. L., Appleby, S. J., Fermin, L. M., Henderson, H. V., Wei, J.,
Wells, D. N., \& Oback, B. (2021). Controlled cytoplast arrest and
morula aggregation enhance development, cryoresilience, and in vivo
survival of cloned sheep embryos. \emph{Cellular Reprogramming},
\emph{23}(1), 14--25.
doi:\href{https://doi.org/10.1089/cell.2020.0078}{10.1089/cell.2020.0078}

\leavevmode\hypertarget{ref-CRAN}{}%
R Core Team. (2021). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-wickham2019welcome}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., et al. (2019). Welcome to the tidyverse.
\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}
